---
title: "FML Assignment 3 Naive Bayes"
author: "Dharani"
date: "2022-10-26"
output: html_document
---
#load the required library
```{r}
library("dplyr")
library("tidyr")
library("ggplot2")
library("ROCR")
library("rpart")
library("rpart.plot")
library("caret")
library("randomForest")
library("tidyverse")
library("tm")
library("SnowballC")
library("softImpute")
library("glmnet")
library("Hmisc")
library('tinytex')
library('GGally')
library('gplots')
library('FNN')
library("caTools")
library("ggpubr")
library("reshape2")
library("e1071")
library("melt")
```
#data from a CSV file, converts relevant columns to factors, sets a seed for reproducibility, and splits the data into training and testing sets.
```{r}
rm(list=ls())
bank = read.csv("C:/Users/DHARANI/OneDrive/Desktop/FML/Assignments/UniversalBank.csv")
bank$Personal.Loan = as.factor(bank$Personal.Loan)
bank$Online = as.factor(bank$Online)
bank$CreditCard = as.factor(bank$CreditCard)
set.seed(1)
train.index <- sample(row.names(bank), 0.6*dim(bank)[1])  
test.index <- setdiff(row.names(bank), train.index) 
train.df <- bank[train.index, ]
test.df <- bank[test.index, ]
train <- bank[train.index, ]
test = bank[train.index,]
```

#reshaping the data to understand relationships between variables.
```{r message=FALSE, warning=FALSE}
melted.bank = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")
recast.bank=dcast(melted.bank,CreditCard+Personal.Loan~Online)
recast.bank[,c(1:2,14)]
```
#reshapes data for understanding relationships.
```{r message=FALSE}
melted.bankc1 = melt(train,id=c("Personal.Loan"),variable = "Online")
melted.bankc2 = melt(train,id=c("CreditCard"),variable = "Online")
recast.bankc1=dcast(melted.bankc1,Personal.Loan~Online)
recast.bankc2=dcast(melted.bankc2,CreditCard~Online)
```
#extracting specific columns from reshaped data.
```{r}
Loanline=recast.bankc1[,c(1,13)]
LoanCC = recast.bankc2[,c(1,14)]
Loanline
LoanCC
```
#generating frequency tables for specific columns in the training data.
```{r}
table(train[,c(14,10)])
table(train[,c(13,10)])
table(train[,c(10)])
```
#calculating probabilities related to specific events.
```{r}
probability1<-77/(77+198)
probability1
```

```{r}
probability2<-166/(166+109)
probability2
```

```{r}
probability3<-275/(275+2725)
probability3

```

```{r}
probability4<-801/(801+1924)
probability4
```

```{r}
probability5<-1588/(1588+1137)
probability5
```

```{r}
probability6<-2725/(2725+275)
probability6
```

```{r}
(probability1*probability2*probability3)/((probability1*probability2*probability3)+(probability4*probability5*probability6))

```
#data for a Naive Bayes model, splitting it into training and testing sets, and then fits a Naive Bayes classifier on the training data.
```{r}
naive.train = train.df[,c(10,13:14)]
naive.test = test.df[,c(10,13:14)]
naivebayes = naiveBayes(Personal.Loan~.,data=naive.train)
naivebayes
```