---
title: "Dharani FML Assignment 1"
Author: Dharani
date: "2024-02-04"
output:
  pdf_document: default
  html_document: default
Source: "Kaggle (https://www.kaggle.com/datasets/eugeniyosetrov/airline-delays/code)"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

#This is airline delay data set which is taken from Kaggle and imported
```{r}
airline_data <- read.csv("C:/Users/DHARANI/OneDrive/Desktop/FML/airline_delay.csv")
```

#Print out descriptive statistics for a selection of quantitative and categorical variables.

```{r}
summary(airline_data)
head(airline_data)
tail(airline_data)
```
#Transform at least one variable.

#Transformed arr_diverted column was originally in integer form; however, it is now being transformed into characters by changing 0 and 1 to represent diverted and non-diverted.
```{r}
airline_data$arr_diverted <- ifelse(airline_data$arr_diverted == 1, "diverted", "not_diverted")
airline_data <- na.omit(airline_data)
as.tibble(airline_data)
```

#Plot at least one quantitative variable, and one scatterplot.

#Here is the scattered plot of the overall delay time versus the total arrival by time after transformation.
```{r}
##data(mdeaths)
##head(mdeaths)

scatter_plot <- ggplot(airline_data, aes(x = arr_flights, y = arr_delay)) +
  geom_point(  color = "purple") +
  labs(title = "Total delay time vs. Total arrival by Airport",
       x = "Total Arrivals",
       y = "Total Delay Time")
scatter_plot
```

